Reproducible Research Project 1 - Elizabeth Lundeen
=============================

**Step 1: Code for reading in the data set and/or processing the data (exploring the data).**

```{r, echo=TRUE}
setwd("C:/Users/yxj4/OneDrive - CDC/+My_Documents/CDC/1 DDT/Data Modernization/Coursera/Hopkins Data Science Specialization/5 Reproducible Research/3 Course Project 1")
activity <- read.csv("activity.csv")
head(activity,10)
str(activity)

```

**Step 2: Show the distribution of total number of steps per day and calculate the mean and median total number of steps per day.**

* Calculate the total number of steps taken per day.    
* Make a histogram of the total number of steps taken each day.  
* Calculate and report the mean and median of the total number of steps taken per day.  

Answer: the mean number of steps per day is 10,766.19 and the median number of steps per day is 10,765.

```{r, echo=TRUE}
library(ggplot2)
activity$date <- as.Date(activity$date)
totalsteps <- aggregate(steps ~ date, activity, sum)
qplot(steps, data = totalsteps, bins=8, xlab = "Total Number of Steps", ylab = "Frequency, Number of Days", main = "Histogram of Total Number of Steps Per Day")

meansteps <- mean(totalsteps$steps, na.rm=TRUE)
mediansteps <- median(totalsteps$steps, na.rm=TRUE)

meansteps
mediansteps

```

**Step 3: Time Series Plot to Show the Average Daily Activity Pattern Per 5-Minute Interval**

* Make a time series plot (type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).  
* Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?  

Answer: Interval 835 contains the maximum number of steps (206.1698 steps)

```{r, echo=TRUE}
intervalsteps <- aggregate(steps ~ interval, data = activity, mean)
str(intervalsteps)
plot(intervalsteps,type="l")

intervalsteps[which.max(intervalsteps$steps),]
```

**Step 4: Imputing Missing Values**
* Calculate and report the total number of missing values in the data set (i.e. the total number of rows with NAs).    
* Impute the missing values in the dataset by using the mean for that 5-minute interval.  
* Create a new data set that is equal to the original data set but with the missing data filled in.  
* Make a histogram of the total number of steps taken each day.    
* Calculate and report the mean and median total number of steps taken per day.  
* Do these values differ from the estimates from the first part of the assignment?   
* What is the impact of imputing missing data on the estimates of the total daily number of steps?  

Answer:  
* Total number of missing values in the data set (steps variable): 2304  
* Mean total number of steps taken per day: 10766.19  
* Median total number of steps taken per day: 10765   
* The mean and median don't differ from the first part of the assignment however the total daily number of steps is larger.  


```{r, echo=TRUE}
sum(is.na(activity$steps))

activity2 <- activity
miss <- is.na(activity2$steps)
avg_interval <- tapply(activity2$steps, activity2$interval, mean, na.rm=TRUE, simplify=TRUE)
activity2$steps[miss] <- avg_interval[as.character(activity2$interval[miss])]

sum(is.na(activity2$steps))

activity2$date <- as.Date(activity2$date)
totalsteps2 <- aggregate(steps ~ date, activity2, sum)
qplot(steps, data = totalsteps2, bins=11, xlab = "Total Number of Steps", ylab = "Frequency, Number of Days", main = "Histogram of Total Number of Steps Per Day (Replace NAs)")

meansteps2 <- mean(totalsteps2$steps, na.rm=TRUE)
mediansteps2 <- median(totalsteps2$steps, na.rm=TRUE)

meansteps
mediansteps
```

**Step 5: Determine if there are differences in activity patterns between weekdays and weekends**

```{r, echo=TRUE}
library(lattice)
week_days <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
activity2$dayofweek <- factor((weekdays(activity2$date) %in% week_days), levels=c(FALSE, TRUE), labels=c('weekend', 'weekday'))
table(activity2$dayofweek)

averagesteps <- aggregate(steps ~ interval+dayofweek, data = activity2, mean)
str(averagesteps)
xyplot(steps ~ interval | dayofweek, data = averagesteps, type="l", layout = c(1, 2), main = "Comparison of Average Steps by Interval Weekdays and Weekends")
```

